---
title: | 
 | Time Series Analysis
 | Border Crossing
author: "Rodrigo Meneses Orellana"
date: "27 de octubre de 2019"
output: pdf_document
header-includes: 
 \usepackage{float}
 \usepackage{longtable}
 \usepackage{xcolor}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
list.of.packages <- c("ggplot2","kableExtra","gridExtra","forecast","tseries","ggfortify")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
require(ggplot2)
require(kableExtra)
require(gridExtra)
library(forecast)
library(tseries)
library(ggfortify)

datos<-read.csv(choose.files(caption = "Seleccione Border Crossing Data"))
```

```{r,echo=0,message=0}
# Limpieza y transformación de los datos: es necesario desagregar las fechas y las posiciones geográficas.

df1<-strsplit(as.character(datos$Date),"/")
meses<-c()
for(i in 1:length(df1)){
  meses[i]<-df1[[i]][1]
}

meses.abb<-month.abb[as.numeric(meses)]

years<-c()
years2<-c()
for(i in 1:length(df1)){
  years[i]<-strsplit(df1[[i]][3]," ")
  years2[i]<-years[[i]][1]
}

# Los datos observados corresponden a información mensual desde 1996 hasta marzo del 2019.

posiciones<-strsplit(substr(datos$Location,8,nchar(as.character(datos$Location))-1)," ")

long<-c()
lat<-c()
for(i in 1:length(posiciones)){
  long[i]<-posiciones[[i]][1]
  lat[i]<-posiciones[[i]][2]
}

datos2<-cbind(datos[,c(1,2,3,4,6,7)],month=meses.abb,year=as.numeric(years2),long,lat)
```


# Introduction

An study of the behavior of border crossing data since January 1996 until March 2019 based on time series.

# Data review

**Taken from GP databases.**

Information related with the border crossing between US-Mexico and US-Canada grouped by the most common ways.

In order to present a not exhaustive study based on time series, currently unnecessary information will be suppressed, leading to the next variables (the study could be expanded).

Preview:

```{r,fig.align='c',fig.pos = 'H', out.width='100%',echo=0}
kable(head(datos2[,c("year","month","Measure","Value")]), format = "latex", 
      caption = "Database sample",
      booktabs=TRUE, longtable=TRUE) %>%
  kable_styling(latex_options = "HOLD_position")
```

Data will be aggregated by year and month in order to form the time series. Aggregation leads to a stacked structure where each rectangle is an aggregation by Year/Month and Measure.

```{r,fig.align='c',fig.pos = 'H', out.width='100%',echo=0,fig.cap="Data aggregation"}
agg1<-aggregate(datos2$Value,by=list(datos2$year,datos2$month,datos2$Measure),sum)
colnames(agg1)<-c("year","month","Measure","Value")
p1<-ggplot(agg1)+geom_col(aes(x=year,y=Value,fill=Measure))+theme_bw()+labs(x="Year",y="By Year")+ theme(legend.position = "top") 
p2<-ggplot(agg1)+geom_col(aes(x=month,y=Value,fill=Measure))+theme_bw()+labs(x="Year",y="By Month")+ theme(legend.position = "none") 

grid.arrange(p1,p2)
```

# Constructing and analyzing the Time Series

Aggregation by month and year leads to the time series data needed.

Then, it's easy to construct the time series.

```{r,fig.align='c',fig.pos = 'H', out.width='100%',echo=0}
datosagg<-aggregate(datos2$Value,by=list(datos2$year,datos2$month),sum)
colnames(datosagg)<-c("year","month","value")

datats<-ts(datosagg[,c("year","month","value")],start = c(1996,1),frequency = 12)

kable(head(datats), format = "latex", 
      caption = "Time series data sample",
      booktabs=TRUE, longtable=TRUE) %>%
  kable_styling(latex_options = "HOLD_position")

datats<-datats[,-c(1,2)]
```

Then, here's a visualization of the time series.

```{r,fig.align='c',fig.pos = 'H', out.width='100%',echo=0,fig.cap="Time series"}
autoplot(datats,ts.colour = "blue",ts.linetype = "dashed")+theme_bw()
```

The time series seems to be stationary about the mean and variance. As a visual prove the autocorrelation function will be plotted.

```{r,fig.align='c',fig.pos = 'H', out.width='100%',echo=0,fig.cap="Autocorrelation function"}
autoplot(acf(datats, plot = FALSE))+theme_bw()
```

Looking at the autocorrelation function is possible to say that the time series is stationary.

Next, each component of the time series will be studied.

```{r,fig.align='c',fig.pos = 'H', out.width='100%',echo=0,fig.cap="Component disaggregation"}
autoplot(stl(datats, s.window = "periodic"), ts.colour = "blue")+theme_bw()
```

Given the component analysis, stationarity is fairly acceptable. Time series could be modeled by an ARIMA model. It's necessary to use a test to prove the stationarity. Then, ADF and KPSS tests will be applied where the null hypothesis is: **time series is non-stationary**.

```{r,fig.align='c',fig.pos = 'H', out.width='100%',echo=0,message=0,warning=0}
adf<-adf.test(datats)
kpss<-kpss.test(datats)
a<-data.frame(ADF=adf$p.value,KPSS=kpss$p.value)

kable(a, format = "latex", 
      caption = "ADF and KPSS tests for non-stationarity",
      booktabs=TRUE, longtable=TRUE) %>%
  kable_styling(latex_options = "HOLD_position")

```

Then, non-stationarity is rejected, so an ARIMA model will be fitted.

```{r,echo=0,message=0}
arima1<-auto.arima(datats, stepwise = FALSE, approximation = FALSE)
```

Residuals need to be independent (just white noise) and not to be normally distributed, therefore, is necessary to apply some tests.

```{r,fig.align='c',fig.pos = 'H', out.width='100%',echo=0}
bp<-Box.test(arima1$residuals)
lb<-Box.test(arima1$residuals,type = "Ljung-Box")

b<-data.frame(`Box-Pierce`=bp$p.value,`Ljung-Box`=lb$p.value)

kable(b, format = "latex", 
      caption = "Residual independence tests",
      booktabs=TRUE, longtable=TRUE) %>%
  kable_styling(latex_options = "HOLD_position")
```

Then, **residuals aren't correlated**.

```{r,fig.align='c',fig.pos = 'H', out.width='100%',echo=0}
jb<-jarque.bera.test(arima1$residuals)
sw<-shapiro.test(arima1$residuals)

c<-data.frame(`Jarque-Bera`=jb$p.value,`Shapiro-Wilk`=sw$p.value)

kable(c, format = "latex", 
      caption = "Normal distribution tests for residuals",
      booktabs=TRUE, longtable=TRUE) %>%
  kable_styling(latex_options = "HOLD_position")
```

Then, **residuals aren't normally distributed**.

Finally, the model is ready to be deployed, some predictions will be made.

```{r,fig.align='c',fig.pos = 'H', out.width='100%',echo=0,fig.cap="Forecasting"}
forecast1<-forecast(arima1,level = c(95),h=50)
autoplot(forecast1)+theme_bw()
```
